### 2017-08-28 paper/paper.tex
Working on poster

\def\papertitle{On the design and use of once-differentiable high dynamic
resolution atoms for the Distribution Derivative Method}
\def\paperauthorA{Nicholas Esterer}
\def\paperauthorB{Philippe Depalle}

% Authors' affiliations have to be set below

%------------------------------------------------------------------------------------------
\documentclass[twoside,a4paper]{article}
\usepackage{dafx_17}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{euscript}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{ifpdf}

\usepackage[english]{babel}
\usepackage{caption}
\usepackage{subfig} % or can use subcaption package
\usepackage{color}
\usepackage{cite}
\usepackage{float}

\setcounter{page}{1}
\ninept

\usepackage{hyphenat}
\hyphenation{spher-oidal}
%\hyphenation{amp-li-tude}
%\hyphenation{mod-u-lat-ed}

\usepackage{times}
% Saves a lot of ouptut space in PDF... after conversion with the distiller
% Delete if you cannot get PS fonts working on your system.

% pdf-tex settings: detect automatically if run by latex or pdflatex
\newif\ifpdf
\ifx\pdfoutput\relax
\else
   \ifcase\pdfoutput
      \pdffalse
   \else
      \pdftrue
\fi

\ifpdf % compiling with pdflatex
  \usepackage[pdftex,
    pdftitle={\papertitle},
    pdfauthor={\paperauthorA, \paperauthorB},
    colorlinks=false, % links are activated as colror boxes instead of color text
    bookmarksnumbered, % use section numbers with bookmarks
    pdfstartview=XYZ % start with zoom=100% instead of full screen; especially useful if working with a big screen :-)
  ]{hyperref}
  \pdfcompresslevel=9
  \usepackage[pdftex]{graphicx}
  \usepackage[figure,table]{hypcap}
\else % compiling with latex
  \usepackage[dvips]{epsfig,graphicx}
  \usepackage[dvips,
    colorlinks=false, % no color links
    bookmarksnumbered, % use section numbers with bookmarks
    pdfstartview=XYZ % start with zoom=100% instead of full screen
  ]{hyperref}
  % hyperrefs are active in the pdf file after conversion
  \usepackage[figure,table]{hypcap}
\fi

% Flowchart drawing stuff
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows, positioning, calc}
\tikzstyle{startstop} = [rectangle,draw=black,text centered]%, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
\tikzstyle{io} = [trapezium, trapezium left angle=70, trapezium right angle=110, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=blue!30]
\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=0cm, text
centered, draw=black, text width=3cm]
\tikzstyle{emptybox} = []
\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]
\tikzstyle{arrow} = [thick,->,>=stealth]
\tikzstyle{noarrow} = [thick,-,>=stealth]

\title{\papertitle}

\twoaffiliations{
\paperauthorA \,}
{\href{https://mt.music.mcgill.ca/spcl}{SPCL\sthanks{Sound Processing and Control Laboratory}} \\ McGill University\\ Montreal, Quebec, Canada\\
{\tt \href{mailto:nicholas.esterer@mail.mcgill.ca}{nicholas.esterer@mail.mcgill.ca}}
}
{\paperauthorB \,}
{\href{https://mt.music.mcgill.ca/spcl}{SPCL} \\ McGill University\\ Montreal,
Quebec, Canada\\ {\tt \href{mailto:philippe.depalle@mcgill.ca}{philippe.depalle@mcgill.ca}}
}

\input{ddm_snr_win_comp_defs.txt}
\input{comp_offset_chirp_est_err_defs.txt}

\begin{document}
% more pdf-tex settings:
\ifpdf % used graphic file format for pdflatex
  \DeclareGraphicsExtensions{.png,.jpg,.pdf}
\else  % used graphic file format for latex
  \DeclareGraphicsExtensions{.eps}
\fi

\maketitle

\begin{abstract}
    The accuracy of the Distribution Derivative Method (DDM)
    \cite{betser2009sinusoidal} is evaluated on mixtures of chirp signals. It is
    shown that accurate estimation can be obtained when the sets of atoms for
    which the inner product is large are disjoint.  This amounts to designing
    atoms with windows whose Fourier transform exhibits low sidelobes but which
    are once-differentiable in the time-domain. A technique for designing
    once-differentiable approximations to windows is presented and the accuracy
    of these windows in estimating the parameters of sinusoidal chirps in mixture is
    evaluated.
\end{abstract}

\section{Introduction}
\label{sec:intro}
Additive synthesis using a sum of sinusoids plus noise is a powerful model for
representing audio \cite{serra1989system}, allowing for the easy implementation
of many manipulations such as time-stretching \cite{marchand2004enhanced} and
timbre-morphing \cite{haken2007beyond}.
%
In these papers,
\cite{serra1989system,marchand2004enhanced,haken2007beyond} the
phase evolution of the sinusoid is assumed linear over the analysis frame, only
the phase and frequency of the sinusoids at these analysis points are used to
fit a plausible phase function after some the analysis points are connected to
form a partial \cite{mcaulay1986speech}. 
%
Recently, there has
been interest in using higher-order phase functions \cite{xuepiecewise} as the estimation of
their parameters has been made possible by a new set of techniques of only
moderate computational complexity using signal derivatives \cite{hamilton2012unified}.
%
The use
of higher-order phase models allows for accurate description of highly modulated
signals, for
example in the analysis of birdsong \cite{stowell2013improved}.
%
The frequency
modulation information has also been used in the
regularization of mathematical programs for audio source separation \cite{creager2016musicalsource}.

The sinusoidal model approximating signal $s$ typically considered is
%
\begin{equation}
    \label{eq:polyphaseexp}
    \tilde{s}(t) = \exp(a_0 + \sum_{q=1}^{Q} a_{q} t^q) + \eta(t)
\end{equation}
where $\tilde{s}$ is the approximating signal, $t$ the variable of time, the
$a_{q} \in \mathbb{C}$ coefficients of the argument's polynomial, and $\eta(t)$
white Gaussian noise.
%
Although this technique can be extended to describe a single sinusoid of
arbitrary complexity simply by increasing $Q$, it remains essential to consider
signals featuring a sum of $P$ such components, whether they represent the harmonic structure of a musical sound or the union of partials resulting from a mixture of multiple signal sources (e.g., recordings of multiple speakers or performers),  i.e., 
%
\begin{equation}
    x(t) = \sum_{p=1}^{P} x_{p}(t) + \eta(t)
\end{equation}
%
with
%
\begin{equation}
    \label{eq:polyphaseexpmix}
    x_{p}(t) = \exp(a_{p,0} + \sum_{q=1}^{Q} a_{p,q} t^q)
\end{equation}
%
As regards the design and evaluation of signal-derivatives analysis techniques,
previous work has generally assumed signals containing a single component, i.e.,
$P$ = 1 or assumed the influence of other components to be negligible. Later we
will refine when this assumption can be made. In \cite{hamilton2012comparisons}
the authors provide a comprehensive evaluation of various signal-derivatives
analysis methods applied to a single-component signal. In
\cite{robel2002estimating} the extent to which two components in mixture can
corrupt estimations of the frequency slope ($\Im\{a_{0,2}\}$ and
$\Im\{a_{1,2}\}$) is investigated in the context of the reassignment method, one
of the signal-derivatives techniques, but the corruption of the other
parameters is not considered.

In this paper, we revisit the quality of signal-derivatives estimation of
\textit{all} the $a_{q}$ when analyzing a \textit{mixture} of components. We focus on
the DDM \cite{betser2009sinusoidal} analysis method for its convenience as it
can simply be considered as an atomic decomposition (see
Sec.~\ref{sec:estimating_aj}), and does not require computing derivatives
of the signal to be analysed.

The DDM does, however, require a once-differentiable analysis window. As we are
interested in windows with lower sidelobes in order to better estimate
parameters of sinusoidal chirp signals in mixture, we seek windows that
combine these two properties. For this, a technique to design
once-differentiable approximations to arbitrary symmetrical windows is proposed
and presented along with a design example for a high-performance window. Finally
we evaluate the performance of various once-differentiable windows in estimating
the parameters $a_{q}$.

\section{Estimating the parameters \lowercase{$a_{q}$}}
\label{sec:estimating_aj}
We will now show briefly how the DDM can be used to estimate the $a_{q}$. Based on
the theory of distributions \cite{schwartz1959theorie}, the DDM
makes use of ``test
functions'' or atoms $\psi$. These atoms must be once differentiable with
respect to time variable $t$ and be non-zero only on a finite interval
$[-\frac{L_{t}}{2},\frac{L_{t}}{2}]$. First, we define the inner product
%
\begin{equation}
    \label{eq:ddm:inner:prod:def}
    \left\langle x , \psi \right\rangle = 
    \int_{-\infty}^{\infty}x(t)\overline{\psi}(t)dt
\end{equation}
%
and the operator 
%
\begin{equation}
\mathcal{T}^{\alpha} : (\mathcal{T}^{\alpha}x)(t) = t^{\alpha}x(t)
\end{equation}
%
Consider the weighted signal
%
\begin{equation}
    f(t) = x(t) \overline{\psi}(t)
\end{equation}
%
differentiating with respect to $t$ we obtain
%
\begin{multline}
    \label{eq:ddm:weighted:sig:derivative}
    \frac{df}{dt}(t) = 
    \frac{dx}{dt}(t)\overline{\psi}(t)
    + x(t)\frac{d\overline{\psi}}{dt}(t) = \\
    \left( \sum_{q=1}^{Q} q a_{q} t^{q-1} \right) x(t)\overline{\psi}(t)
    + x(t)\frac{d\overline{\psi}}{dt}(t)
\end{multline}
%
Because $\psi$ is zero outside of the interval $[-\frac{L_{t}}{2},\frac{L_{t}}{2}]$, integrating
$\frac{df}{dt}(t)$ we obtain
%
\begin{multline}
    \int_{-\infty}^{\infty}\frac{df}{dt}(t)dt = \\
    \sum_{q=1}^{Q} q a_{q} \int_{-\frac{L_{t}}{2}}^{\frac{L_{t}}{2}} t^{q-1} x(t) \overline{\psi}(t) dt
    + \left\langle x, \frac{d\overline{\psi}}{dt} \right\rangle = 0
\end{multline}
%
or, using the operator $\mathcal{T}^{\alpha}$,
%
\begin{equation} 
    \sum_{q=1}^{Q} q a_{q} 
    \left\langle \mathcal{T}^{q-1} x , \overline{\psi} \right\rangle
    = -\left\langle x, \frac{d\overline{\psi}}{dt} \right\rangle
\end{equation}
%

Estimating coefficients $a_{q}$, ${1 < q \leq Q}$, simply requires $R$ atoms
$\psi_{r}$ with $R \geq Q$ to solve the linear system of equations
\begin{equation}
    \label{eq:ddmsyseq}
    \sum_{q=1}^{Q} q a_{q} 
    \left\langle \mathcal{T}^{q-1} x , \overline{\psi}_{r} \right\rangle
    = -\left\langle x, \frac{d\overline{\psi}_{r}}{dt} \right\rangle
\end{equation}
for $1 \leq r \leq R$.

To estimate $a_0$ we rewrite the signal we are analysing as
\begin{equation}
    x(t) = \exp(a_0) \gamma(t) + \epsilon (t)
\end{equation}
where $\epsilon (t)$ is the error signal, the part of the signal that is not explained
by our model, and $\gamma (t)$ is the part of the signal
whose coefficients have already been estimated, i.e.,
\begin{equation}
    \gamma(t) = \exp \left( \sum_{q=1}^{Q} a_{q} t^{q} \right)
\end{equation}
Computing the inner product $\left\langle x , \gamma \right\rangle$, we have
\begin{equation}
    \left\langle x , \gamma \right\rangle
    =
    \left\langle \exp(a_0) \gamma , \gamma \right\rangle + 
        \left\langle \epsilon , \gamma \right\rangle
\end{equation}
The inner product between $\epsilon$ and $\gamma$ is $0$, by the orthogonality
principle \cite[ch.~12]{kay1993fundamentals}. Furthermore, because $\exp(a_0)$ does not
depend on $t$, we have
\begin{equation}
    \left\langle x , \gamma \right\rangle
    =
    \exp(a_0) \left\langle \gamma , \gamma \right\rangle
\end{equation}
so we can estimate $a_0$ as
\begin{equation}
    \label{eq:ddmesta0}
    a_0 = \log \left( \left\langle x , \gamma \right\rangle \right)
        - \log \left( \left\langle \gamma , \gamma \right\rangle \right)
\end{equation}

As will be seen in subsequent sections, the DDM typically involves taking the
discrete Fourier transform (DFT) of the signal windowed by both an everywhere once-differentiable function of
finite support (e.g., the Hann window) and this function's derivative. A small
subset of atoms corresponding to the peak bins in the DFT are used in
Eq.~\ref{eq:ddmsyseq} to solve for the parameters $a_{q}$.

\section{Estimating the \lowercase{$a_{p,q}$} of $P$ components}

We examine how the mixture model influences the estimation of the $a_{p,q}$ in
Eq.~\ref{eq:polyphaseexpmix}.
Consider a mixture of $P$ components.
If we define the weighted signal sum
%
\begin{equation}
    g(t) = \sum_{p=1}^{P} x_{p}(t) \overline{\psi}(t) = \sum_{p=1}^{P} f_{p}(t)
\end{equation}
%
and substitute $g$ for $f$ in Eq.~\ref{eq:ddm:weighted:sig:derivative} we obtain
%
\begin{multline}
    \label{eq:mixest}
    \sum_{p=1}^{P} \int_{-\frac{L_{t}}{2}}^{\frac{L_{t}}{2}}
    \frac{df_{p}}{dt}(t)dt =
    0
    = \\ 
    \sum_{p=1}^{P} \left(
    \sum_{q=1}^{Q} q a_{p,q} 
    \left\langle \mathcal{T}^{q-1} x_{p} , \overline{\psi} \right\rangle
    + \left\langle x_{p}, \frac{d\overline{\psi}}{dt} \right\rangle \right)
\end{multline}
%
From this we see if $\left\langle \mathcal{T}^{q-1} x_{p} , \overline{\psi}_{r}
\right\rangle$ and $\left\langle x_{p}, \frac{d\overline{\psi_{r}}}{dt} \right\rangle$
are small for all but $p = p^{\ast}$ and a subset of $R$ atoms\footnote{%
The notation $x^{\ast}$ will mean the value of the argument $x$ maximizing or minimizing some
function.
}%
, we can simply estimate the parameters $a_{p^{\ast},q}$ using
\begin{equation}
    \sum_{q=1}^{Q} q a_{{p^{\ast}},q} 
    \left\langle \mathcal{T}^{q-1} x_{p^{\ast}} , \overline{\psi}_{r} \right\rangle
    = -\left\langle x_{p^{\ast}}, \frac{d\overline{\psi}_{r}}{dn} \right\rangle
\end{equation}
for $1 \leq r \leq R$. To compute $a_{p^{\ast},0}$ we simply use
\begin{equation}
    \gamma_{p^{\ast}}(t) = \exp \left( \sum_{q=1}^{Q} a_{p^{\ast},q} t^{q} \right)
\end{equation}
in place of $\gamma$ in Eq.~\ref{eq:ddmesta0}.

\section{Designing the $\psi_{r}$}
\label{sec:designingatoms}
%
\begin{figure*}[t]
    \centerline{\includegraphics[width=7in]{{search_dpw_bw_m}.eps}}
\caption{\label{fig:dpw} Comparing the main-lobe and asymptotic power
spectrum characteristics of the continuous 4-term Nuttall window, the digital
prolate window with $W=0.008$, and the continuous approximation to the digital
prolate window.}
\end{figure*}
%
In practice, an approximation of Eq.~\ref{eq:ddm:inner:prod:def} is evaluated using
the DFT on a signal $x$ that is properly sampled
and so can be evaluated at a finite number of times $nT$ with $n \in [0,N-1]$ and
$T$ the sample period in seconds. In this way, the chosen atoms $\psi_{\omega}(t)$ 
are the products of the elements of the Fourier basis and an appropriately
chosen window $w$ that is once differentiable and finite, i.e.,
%
\begin{equation}
    \label{eq:fourieratom}
    \psi_{\omega}(t) = w(t) \exp(-j \omega t)
\end{equation}
%
Defining $N = \frac{L_{t}}{T}$ and angular frequency at bin $r$ as $\omega_{r} = 2
\pi \frac{r}{N}$, the approximate inner product is then
%
\begin{equation}
    \label{eq:approxinnerprod}
    \left\langle x , \psi_{\omega} \right\rangle \approx 
    \sum_{n=0}^{N-1} x(Tn) w(Tn) \exp(-2 \pi j r \frac{n}{N}) 
\end{equation}
%
i.e., the definition of the DFT of a windowed signal%
\footnote{%
    Notice however that this is an approximation of the inner product and should
    not be interpreted as yielding the Fourier series coefficients of a properly
    sampled signal $x$ periodic in $L_{t}$. This means that other
    evaluations of the inner product that yield more accurate results are
    possible. For example, the analytic solution is possible if $x$ is assumed
    zero outside of $[-\frac{L_{t}}{2},\frac{L_{t}}{2}]$ (the $\psi$ are in
    general analytic).  In this case the samples of $x$ are convolved with the
    appropriate interpolating sinc functions and the integral of this function's
    product with $\psi$ is evaluated.
}. %
The DFT is readily interpreted as a bank of bandpass filters centred at
normalized frequencies%
%\footnote{%
%In this article as in much of the literature on signal processing the index $r$
%of the Fourier basis element corresponding to the frequency $\omega_{r} = 2 \pi
%\frac{r}{N}$ will be referred to as its bin number.
%}
$\frac{r}{N}$ and with frequency response described by the Fourier transform of
modulated $w$ \cite{allen1977unified}. Therefore choosing $\psi$ amounts to a
filter design problem under the constraints that the impulse response of the
filter be differentiable in $t$ and finite. To minimize the influence of all but
one component, granted the components's energy concentrations are sufficiently
separated in frequency, we desire impulse responses whose magnitude response
gives maximum out-of-band rejection or equivalently, windows whose Fourier
transform exhibits the lowest sidelobes.

In all the publications reviewed on the DDM for this paper, the window used was
the Hann window which is once-differentiable everywhere in the time-domain. In
\cite{robel2002estimating}, a publication on the reassignment method, other
windows than the Hann are considered but these windows must be
twice-differentiable.  Nuttall \cite{nuttall1981some} has designed windows with
lower sidelobes than the canonical Hann window which are everywhere at least
once-differentiable. It is also possible to design approximations to arbitrary
symmetrical window functions using harmonically related cosines, as is discussed
in the following section.

\section{Differentiable approximations to windows}
%
\begin{figure*}[ht]
    \centerline{\includegraphics[width=7in]{{ddm_snr_win_comp}.eps}}
\caption{\label{fig:snrwincomp} The estimation variance of random polynomial
    phase sinusoids averaged over $K_{1}=\Ksnr{}$ trials using atoms generated from
    various windows. \textit{C} is the Cram\'{e}r-Rao
lower bound, \textit{N3} and \textit{N4} are the 3- and 4-cosine-term continuous
Nuttall windows, \textit{H} is the Hann window, and \textit{P5} is the continuous
5-cosine-term approximation to a digital prolate window as described in
Sec.~\ref{sec:designexample}.}
\end{figure*}
%
A differentiable approximation to a symmetrical window can be designed in a
straightforward way. In \cite{harris1978use} and \cite{rabiner1970approach} it
is shown how to design optimal windows of length $N$ samples using a linear
combination of $M$ harmonically related cosines
\begin{equation}
    \tilde{w}(n) = \sum_{m=0}^{M-1} b_{m} \cos (2 \pi m \frac{n}{N})
\mathcal{R}(\frac{n}{N})
\end{equation}
where $\mathcal{R}$ is the \textit{rectangle function}. This function is
discontinuous
at $n = \frac{\pm N}{2}$, and therefore not differentiable there, unless
\begin{equation}
\sum_{m=0}^{M-1} b_{m} \cos ( \pm \pi m ) = 0
\end{equation}

Rather than design based on an optimality criterion, such as the height of the
highest sidelobe \cite{rabiner1970approach}, a once-differentiable approximation
to an existing window $w$ is desired. To do this, we choose the $b_{m}$ so that
the window $\tilde{w}$'s squared approximation error to $w$ is minimized while having
$\tilde{w}(\frac{\pm N}{2}) = 0$, i.e. we find the solution $\{ b^{\ast}_{m} \}$ to the
mathematical program
\begin{equation}
    \label{eq:searchcontwinprogram}
    \text{minimize}
    \sum_{n=0}^{N-1} ( w(n) 
        - \sum_{m=0}^{M-1} b_{m} \cos(2 \pi m \frac{n}{N}))^{2}
\end{equation}
\begin{equation}
    \text{subject to} \\
    \sum_{m=0}^{M-1} b_{m} \cos(\pi m ) = 0
\end{equation}
which can be solved using constrained least-squares; a standard numerical
linear algebra routine \cite[p.~585]{golub1996matrix}.

\section{A continuous window design example}
\label{sec:designexample}
%
\begin{figure*}[ht]
    \centerline{\includegraphics[width=7in]{{comp_offset_chirp_est_err_params}.eps}}
\caption{\label{fig:mixwincomp} The mean squared estimation error for each
    parameter in an analysis of two components in mixture. A set of
    $K_{2}=\Koffset{}$ chirps was synthesized and each unique pair used for
    maximum bin differences $0 \leq d < \Doffset{}$, with $d$ varied in
    $\Dstep{}$ bin increments. The signal power ratio
    between components is indicated with colours and the corresponding ratio in decibels is indicated in the
    plot legend. The names indicate the windows used to generate the atoms for
    estimation: \textit{N3} and \textit{N4} are the 3- and 4-cosine-term
    continuous Nuttall windows, \textit{H} is the Hann window, and \textit{P5}
is the continuous 5-cosine-term approximation to a digital prolate window as
described in Sec.~\ref{sec:designexample}.}
\end{figure*}
%
As a design example we show how to create a continuous approximation of a digital
prolate spheroidal window.

Digital prolate spheroidal windows are a parametric approximation to functions
whose Fourier transform's energy is maximized in a given bandwidth
\cite{slepian1978prolate}. These can be tuned to have extremely low sidelobes,
at the expense of main-lobe width.  Differentiation of these window functions
may be possible but is not as straightforward as differentiation of the
sum-of-cosine windows above. Furthermore, the windows do not generally have
end-points equal to 0. In the following we will demonstrate how to approximate
a digital prolate spheroidal window with one that is everywhere at least
once-differentiable.

\begin{table}[]
    \caption{The coefficients of the once-differentiable approximation to a digital prolate
    window designed in Sec.~\ref{sec:designexample}.
    \label{tab:contprolate}}
    \begin{center}
        \begin{tabular}{c c c}
            $b_0$ & $ = $ & 3.128 $\times 10^{-1}$ \\
            $b_1$ & $ = $ & 4.655 $\times 10^{-1}$ \\
            $b_2$ & $ = $ & 1.851 $\times 10^{-1}$ \\
            $b_3$ & $ = $ & 3.446 $\times 10^{-2}$ \\
            $b_4$ & $ = $ & 2.071 $\times 10^{-3}$ 
            %3.128258203861262743 $\times 10^{-1}$ &
            %4.655346501582137142 $\times 10^{-1}$ &
            %1.851027605614040672 $\times 10^{-1}$ &
            %3.446534984178639682 $\times 10^{-2}$ &
            %2.071419052469766937 $\times 10^{-3}$ \\
        \end{tabular}
    \end{center}
\end{table}%

In \cite{verma1996digital} it was shown how to construct digital prolate
spheroidal windows under parameters $N$, the window length in samples, and a
parameter $W$ choosing the (normalized) frequency range in which the proportion
of the main lobe's energy is to be maximized. We chose $N=512$ based on the
window length chosen in \cite{betser2009sinusoidal} for ease of comparison. Its
$W$ parameter's value was chosen by synthesizing windows with $W$ ranging
between $0.005$ and $0.010$ at a resolution of $0.001$. The window with the
closest 3 dB bandwidth to the 4-term Nuttall window was obtained with $W=0.008$.
Its magnitude response is shown in Fig.~\ref{fig:dpw}. We see that this window's
asymptotic falloff is 6 dB per octave and therefore has a discontinuity
somewhere in its domain \cite{nuttall1981some}.

We designed an approximate
window using Eq.~\ref{eq:searchcontwinprogram} for $M$ varying between $2$ and
$N/8$ to find the best approximation to the digital prolate window's main lobe
using a small number of cosines. The $M$ giving the best approximation was $5$.
The magnitude response of the approximation is shown in Fig.~\ref{fig:dpw} and
its coefficients are listed in Tab.~\ref{tab:contprolate}; the temporal shape is
very close to a digital prolate spheroidal window with $W=0.008$ and is
therefore omitted for brevity.

It is seen that a lower highest sidelobe level than the Nuttall and Prolate
windows is obtained by slightly sacrificing the narrowness of the main lobe.
More importantly, in Fig.~\ref{fig:dpw} we observe that the falloff of the
window is 18 dB per octave because it is once-differentiable at all points in
its domain.

\section{The performance of improved windows}

\subsection{Signals with single component}

To compare the average estimation error variance with the
theoretical minimum given by the Cram\'{e}r-Rao bound we synthesized $K_{1}$ random
chirps using Eq.~\ref{eq:polyphaseexp}
with $Q=2$ and parameters chosen from uniform distributions justified in
\cite{betser2009sinusoidal}. The original Hann window,
the windows proposed by Nuttall and the new digital prolate based window were
used to synthesize the atoms as described in Sec.~\ref{sec:designingatoms} and
their estimation error variance was compared (see
Fig.~\ref{fig:snrwincomp}). After performing the DFT to obtain inner products
with the atoms, the three atoms whose inner products were greatest were used in
the estimations, i.e., $R=3$ in Eq.~\ref{eq:ddmsyseq}. The windows with the
lowest sidelobes only give the lowest error variance at very favourable SNRs, at
real-world SNRs the original Hann window still performs best at estimating the
parameters of a single component signal.

\begin{figure}[H]
    \centering
            \begin{tikzpicture}[node distance=0.25cm]
                \node (start) [process] at (0,0) {Synthesize $K_{2}$
                single-component signals.};
                \node (modulate) [process, below = of start] {Modulate so that
                peak bin is at frequency 0 for all signals.};
                \node (setd) [process, below = of modulate] {Set $d=0$.};
                \node (choosepair) [process, below = of setd] {Choose
                    $K_{2}(K_{2}-1)$ pairs of signals.};
                \node (dum1) [emptybox, right = of choosepair] {};
                \node (scaleone) [process, below = of choosepair] {Scale one in
                each pair to give desired power from $\mathcal{S}$ and modulate to peak bin
                according to $d$.};
                \node (addpair) [process, below = of scaleone] {Add each signal
                pair together.};
                \node (tryestimate1) [process, below = of addpair] {For each
                pair, try
                estimating parameters of unmodulated component with atoms at
                bins $\{-1,0,1\}$.};
                \node (tryestimate2) [process, below = of tryestimate1] {For
                each pair, try
                estimating parameters of modulated component with atoms at bins
                $\{d-1,d,d+1\}$.};
                \node (sumerrs) [process, below = of tryestimate2] {Sum estimation errors of
                each parameter and divide by $K_{2}(K_{2}-1)$.};
                \node (incd) [process, below = of sumerrs] {Increment $d$ by
                $\Delta_{d}$.};
                \node (dum2) [emptybox, right = of incd] {};
                \node (dum3) [emptybox, below = of incd] {};
                \node (dum3b) [emptybox, left = of incd] {};
                \node (dum4) [emptybox] at (dum3b |- dum3) {};
%                \node let \p{dum3b}=(dum3b), \p{dum3}=(dum3) in (dum4) [emptybox] at (\x{dum3b},\y{dum3}) {};
                \node (dum5) [emptybox, left = of setd] {};

                \draw [arrow] (start) -- (modulate);
                \draw [arrow] (modulate) -- (setd);
                \draw [arrow] (setd) -- (choosepair);
                \draw [arrow] (choosepair) -- (scaleone);
                \draw [arrow] (scaleone) -- (addpair);
                \draw [arrow] (addpair) -- (tryestimate1);
                \draw [arrow] (tryestimate1) -- (tryestimate2);
                \draw [arrow] (tryestimate2) -- (sumerrs);
                \draw [arrow] (sumerrs) -- (incd);

                \draw [arrow] (incd) -- (dum2.center) -- (dum2.center) -- 
                    node [anchor=west,text width=2cm,minimum width=3cm] {If $d <
                    D$.} (dum1.center) 
                    -- (dum1.center) -- (choosepair);

                \draw [arrow] (incd) -- (dum3.center) -- (dum3.center) --
                (dum4.center) -- (dum4.center) -- 
                node [anchor=east,text width=2cm,minimum width=2cm]
                {If signal power ratios in $\mathcal{S}$ remaining to be
                evaluated.} (dum5.center) --
                 (dum5.center) -- (setd);

            \end{tikzpicture}
    \caption{The evaluation procedure for 2-component
    signals.\label{fig:2cevalflowchart}}
\end{figure}

\subsection{Signals with 2 components}

To evaluate the performance of the various windows when estimating the
parameters of components in mixture we synthesized signals using
Eq.~\ref{eq:polyphaseexpmix} with $P=2$ and $Q=2$ and parameters chosen from the
uniform distributions specified in \cite{betser2009sinusoidal}.  We desired to
see how the accuracy of estimation is influenced by the difference (in bins)
between the locally maximized atoms and the difference in signal power between
the two components. To obtain a set of components from which test signals
exhibiting the desired differences could be constructed, we synthesized a set
$\mathcal{C}$ of $K_{2}$ components for which the energy is maximized in bin 0.
Test signals were obtained by choosing a pair of unique components from this set
and modulating one to give the desired frequency and amplitude difference.  This was carried out as
follows: the atom $r^{\ast}$ for which the inner product was maximized was
determined for each unmixed chirp and the chirp was modulated by $\exp(-2\pi
\frac{r^{\ast} n}{N}j)$ for $0 \leq n < N$ in order to move this maximum
to $r=0$. Then for each desired difference $d$, with $0 \leq d < D$ (for the
evaluation $D = \Doffset{}$), two unique
chirps were selected from $\mathcal{C}$ and one chirp was modulated by
$\exp(2\pi \frac{n d}{N}j)$ for $0 \leq n < N$ in order to give the
desired difference between maxima. This component was also scaled by a constant
to give a desired signal power ratio from set $\mathcal{S}$ with the other component (the power ratios
$\mathcal{S}$ tested were 0 dB and -30 dB). As we assume perfect peak-atom selection for this
evaluation no inner-product maximizing $r^{\ast}$ is chosen, rather atoms with
angular frequencies $\omega=2\pi\frac{\hat{d}}{N}$ for $\hat{d} \in
\{d-1,d,d+1\}$ in Eq.~\ref{eq:fourieratom} (again, $R=3$) were chosen to carry
out the estimation. $d$ was incremented by $\Delta_{d} = \Dstep{}$ and so
$\hat{d}$ was not generally integral valued in this case. The parameters of the
unmodulated component were estimated using angular frequencies
$\omega=2\pi\frac{\hat{d}}{N}$ for $\hat{d} \in\{-1,0,1\}$ in
Eq.~\ref{eq:fourieratom}. The squared
estimation error for each parameter was summed and divided by $K_{2}(K_{2}-1)$
(the number of choices of two unique components) to give the averaged squared
estimation error for each parameter at each difference $d$. The procedure is
summarized in Fig.~\ref{fig:2cevalflowchart}.

The behaviour of the windows when used to analyse mixtures of non-stationary
signals is similar to the behaviour of windows used for harmonic analysis in the
stationary case \cite{harris1978use}; here we obtain further insight into how
the estimation of each coefficient of the polynomial in
Eq.~$\ref{eq:polyphaseexp}$ is influenced by main-lobe width and sidelobe
height and slope. In Fig.~\ref{fig:mixwincomp} we see that there is generally less
estimation error for components having similar signal power. This is to be
expected as there will be less masking of the weaker signal in these scenarios.
The estimation error is large when the atoms containing the most signal energy
for each component are not greatly separated in frequency.  This is due to the
convolution of the Fourier transform of the window with the signal, and agrees
with what was predicted by Eq.~\ref{eq:mixest}: indeed windows with a larger
main lobe exhibit a larger ``radius'' (bandwidth) in which the error of the
parameter estimation will be high.  However, for signals where local
inner-product maxima are from atoms sufficiently separated in frequency, windows
with lower sidelobes are better at attenuating the other component and for these
the estimation error is lowest.

\section{Conclusions}

Motivated by the need to analyse mixtures of frequency- and
am\-pli\-tude-mod\-u\-lat\-ed
sinusoids (Eq.~\ref{eq:polyphaseexpmix}), we have shown that the DDM can be
employed under a single-component
assumption when components have roughly disjoint sets of atoms for which their
inner products take on large values. This indicates the need for windows whose
Fourier transform exhibits low sidelobes. We developed windows whose sidelobes
are minimized while remaining everywhere once-differentiable: a requirement to
generate valid atoms for the DDM. These windows were shown to only improve parameter
estimation of $P=1$ component with argument-polynomial of order $Q=2$ in low
amounts of noise. However, for $P=2$ components of the same order in mixture without
noise, granted the components
exhibited reasonable separation in frequency between the atoms for which the
inner product was maximized, these new windows substantially improved the
estimation of all but the first argument-polynomial coefficient.

Further work should evaluate these windows on sinusoids of different orders,
i.e., $Q \gg 1$. Optimal main-lobe widths for windows should be determined
depending on the separation of local maxima in the power spectrum. It should
also be determined if these windows improve the modeling of real-world acoustic
signals.

\section{Acknowledgments}
This work was partially supported by grant from the Natural Sciences and Engineering Research Council of Canada awarded to Philippe Depalle (RGPIN-262808-2012).

%\newpage
%\nocite{*}
\bibliographystyle{IEEEbib}
\bibliography{paper} % requires file DAFx17_tmpl.bib

%\section{Appendix}

\end{document}