### 2017-09-27 poster/sciposter/dafxposter/references.tex

\documentclass[letterpaper,8pt]{article}
\usepackage[margin=0.5in]{geometry}

\begin{document}
\title{\large\textbf{On the design and use of once-differentiable high dynamic
resolution atoms for the Distribution Derivative Method%
}}
\date{}
\pagestyle{headings}
\markright{On the design and use of once-differentiable high dynamic
resolution atoms for the Distribution Derivative Method%
}
\pagenumbering{gobble}
\bibliographystyle{IEEEbib}
\nocite{*}
%\bibliography{dafx-poster} % requires file DAFx17_tmpl.bib
\bibliography{paper}

\end{document}